---
title: "ANA 515 Assignment 2"
author: "Chester Kayonga"
date: "2024-06-12"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=TRUE}
# Load the dataset using the function read.csv
url <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/nba-draft-2015/historical_projections.csv"
Nba2015draft_data <- read.csv(url)
```

```{r, include=TRUE}
# library for data manipulation
library(dplyr)
```

```{r echo=TRUE}
# Filter the dataset to include only rows where Projected.SPM is greater than or equal to 0.2
filtered_data <- Nba2015draft_data %>%
  filter(Projected.SPM >= 0.2)
# Rename columns for better readability
cleaned_data <- filtered_data %>%
  rename(
    Projected_SPM = Projected.SPM,
    Superstar_Prob = Superstar,
    Starter_Prob = Starter,
    Role_Player_Prob = Role.Player,
    Bust_Prob = Bust
  )

# Display the first few rows of the cleaned dataset
head(cleaned_data)
```
# Characteristics of the Data

This dataframe has `r nrow(cleaned_data)` rows and `r ncol(cleaned_data)` columns. The names of the columns and a brief description of each are in the table below:

```{r, echo=TRUE}
# Load necessary library for table creation
library(knitr) 

# Create a dataframe for column descriptions
column_descriptions <- data.frame(
  Column_Name = c("Player", "College", "Year", "Position", "Height", "Wingspan", 
                  "Projected_SPM", "Superstar_Prob", "Starter_Prob", 
                  "Role_Player_Prob", "Bust_Prob"),
  Description = c("Player's Name", "College attended", "Year of draft", "Position played", 
                  "Height in inches", "Wingspan in inches", 
                  "Projected Statistical Plus/Minus", "Probability of becoming a Superstar", 
                  "Probability of becoming a Starter", "Probability of becoming a Role Player", 
                  "Probability of becoming a Bust")
)

# Display the table
kable(column_descriptions, caption = "Column Names and Descriptions")
```

# Summary Statistics

We will generate summary statistics for three columns: `Projected_SPM`, `Superstar_Prob`, and `Starter_Prob`.

# Subsetting the Data
```{r}
# Subset the dataset to include only the columns of interest
data_pick3 <- select(cleaned_data, Projected_SPM, Superstar_Prob, Starter_Prob)
```

# Summary of the Subset

```{r, include=TRUE}
# Produce a summary of the subset
Summarytable <- summary(data_pick3) # Creates the summary
Summarytable # Prints the summary in your output
```
## Data Visualizations 

## Histogram of superstar probability
```{r, echo=TRUE}
# Load necessary library for data visualization
library(ggplot2)

# Create a histogram of Superstar_Prob
ggplot(cleaned_data, aes(x = Superstar_Prob)) +
  geom_histogram(binwidth = 0.05, fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Histogram of Superstar Probability",
       x = "Superstar Probability",
       y = "Frequency") +
  theme_minimal()
```

##Scatterplot of Projected SPM vs. Scatterplot
```{r, echo=TRUE}
# Create a scatter plot of Projected_SPM vs. Starter_Prob
ggplot(cleaned_data, aes(x = Projected_SPM, y = Starter_Prob)) +
  geom_point() +
  labs(title = "Scatter Plot of Projected SPM vs. Starter Probability",
       x = "Projected SPM",
       y = "Starter Probability") +
  theme_minimal()
```

